```{r}
library(ISLR2)
library(ggplot2)
# install.packages("GGally")
library(GGally)
library(gridExtra)
library(reshape2)
data(College)
# head(College)

ggpairs(College) # makes pair plot

 #histogram for Applications feature
p1<-ggplot(College,aes(x=Apps)) +geom_histogram(bins=30)+ggtitle("Applications")
 #histogram for Enrollment feature
p2<-ggplot(College, aes(x=Enroll))+ geom_histogram(bins=30)+ggtitle("Enrollment")
 #histogram for Graduation Date feature
p3<-ggplot(College, aes(x=Grad.Rate))+geom_histogram(bins=30)+ ggtitle("Graduation Rate")
 
grid.arrange(p1,p2,p3,ncol=3)

#correlation matrix
cor_matrix <- cor(College[, sapply(College, is.numeric)])
#  heatmap
cor_melt <- melt(cor_matrix)
ggplot(data=cor_melt, aes(x=Var1, y=Var2, fill=value))+geom_tile(color = "white")+
    scale_fill_gradient2(low = "blue", high="red", mid = "white", midpoint =0, limit=c(-1,1),
    space="Lab",name="Pearson\nCorrelation") +theme_minimal() + 
    theme(axis.text.x = element_text(angle = 45, vjust = 1, size= 12, hjust = 1),
    axis.text.y=element_text(size = 12)) +labs(x= '',y='', title='Correlation Matrix Heatmap') +
    coord_fixed()
```

Scaling or transformation might be necessary for the given data set. If the variables are of different scales and analysis like K-means or PCA is used we will get a wrong output as these algorithm assume the variables to be on the same scale. We should use transformations like square root or log. to normalize the distribution. This won't causxe any errors when using ML algotihms or something else.

```{r}
library(ggplot2)
library(GGally)
library(ISLR2)
library(gridExtra) 
data(College)

College$log_Apps<-log(College$Apps +1) #log transformation, add 1 to 0 values
College$sqrt_Enroll <-sqrt(College$Enroll) # square root
College$sqrt_Grad_Rate <-sqrt(College$Grad.Rate) #square root

#histogram
p1<-ggplot(College,aes(x=log_Apps)) +geom_histogram(bins=30)+ ggtitle("Log of Applications") 
p2<-ggplot(College, aes(x=sqrt_Enroll))+ geom_histogram(bins=30) +ggtitle("Square Root of Enrollment")
p3 <-ggplot(College, aes(x=sqrt_Grad_Rate)) + geom_histogram(bins=30) +
  ggtitle("Square Root of Graduation Rate") 

#arrange plot
grid.arrange(p1, p2,p3,ncol=3)

#pair plot for transformed var.
selected_cols <-c("log_Apps", "sqrt_Enroll","sqrt_Grad_Rate", "Outstate","Room.Board","PhD")
ggpairs(College[, selected_cols])
```

The results after the transformation show that the data distribution have become normal, i.e. reduced skewness. Gives better analysis and clearer insight, for example after the transformation the three variables shown have a more predictable pattern as we are not considering the extreme values anymore.

```{r}
library(ISLR2)
data(College)

#total UG students and their median
College$total_undergrads<- College$F.Undergrad+College$P.Undergrad
median_size <- median(College$total_undergrads)

#subset data into large and small
large <-College[College$total_undergrads > median_size,]
small<-College[College$total_undergrads <=median_size, ]

##sort alphabetically by college name
large<-large[order(rownames(large)), ]
small<-small[order(rownames(small)), ]

#save as txt files
write.table(large,"large_colleges.txt", sep="\t", row.names = TRUE,col.names = NA)
write.table(small,"small_colleges.txt",sep="\t",row.names = TRUE,col.names = NA)
```

```{r}
library(ISLR2)

#loading large and small txt data frames
large <-read.table("large_colleges.txt",sep="\t",header= TRUE)
small<-read.table("small_colleges.txt",sep="\t",header= TRUE)

#median of top 25% for large and small df
median_large_top25=median(large$Top25perc)
median_small_top25=median(small$Top25perc)

#drop rows for top25perc < median
filtered_large = large[large$Top25perc >= median_large_top25, ]
filtered_small = small[small$Top25perc >= median_small_top25, ]

##save
write.table(filtered_large,"filtered_large_colleges.txt",sep="\t",row.names =TRUE,col.names= NA)
write.table(filtered_small,"filtered_small_colleges.txt",sep="\t",row.names =TRUE,col.names= NA)

```

```{r}
filtered_large<-read.table("filtered_large_colleges.txt",sep= "\t",header = TRUE, check.names =FALSE)
filtered_small<-read.table("filtered_small_colleges.txt",sep= "\t",header = TRUE, check.names =FALSE)

#list containing of datasets
college_datasets<- list(
  LargeColleges=filtered_large,
  SmallColleges=filtered_small
)

save(college_datasets,file="CollegeDatasets.RData")
```